---
title: "Assignment 4"
author: "Morgan Baccus"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1
```{r}
# Install and import "nycflights3"

library(nycflights13)
library(tidyverse)
library(dplyr)
library(ggplot2)

```

Problem 1 part a
```{r nycflights13}
# Filter the dataset to display tail number, year, month, day, hour, origin,
# and humidity for all flights heading to TPA after 12pm on Nov. 1, 2013

weather2 <- weather %>%
  select(origin, year, month, day, hour, humid) %>%
  filter(hour >= 12, month == 11, day == 1, year == 2013)

flights_humidity <- flights %>%
 select(year, month, day, hour, origin, dest, tailnum) %>%
  filter(dest == 'TPA', hour >= 12, month == 11, day == 1, year == 2013)
  
flights_humidity %>%
  left_join(weather2, flights_humidity, by= 'hour') %>%
  filter(origin.x == origin.y) %>%
  select(-dest)

```

Problem 1 part b
Explain the difference between the following two joins:

  (1) anti_join(flights, airports, by = c("dest" = "faa"))
  (2) anti_join(airports, flights, by = c("faa" = "dest"))


1 will return any flights that had a destination not in the list of faa destinations.
2 will return any airport in the list of faa destinations that wasn't a destination of any flight in the dataset.

Both joins are looking for matches between dest and faa, but 1 will return flights minus the matches and 2 will return airports minus the matches.





Problem 2 part c
```{r}
# Filter flights to show planes that have flown 100 or more flights

# flights_100 is a list of all the tail numbers that have flown at least 100 flights
flights_100 <- flights %>%
  group_by(tailnum) %>%
  summarise(flight_count = n()) %>%
  filter(flight_count >= 100)

# flight_100_info is a copy of flights,
#but only contains flights that have tail numbers listed in flights_100
flights_100_info <- flights %>%
  semi_join(flights_100)

print(flights_100_info, n=20)
```

Problem 1 part d
```{r}
# What weather conditions make it more likely to see a delay?
# I'm going to use dep_delay to measure the delays
# I'm going to analyze the weather by looking at temp, wind_speed, and precip

flight_weather <- flights %>%
  full_join(weather) %>%
  select(dep_delay, temp, wind_speed, precip) %>%
  filter(dep_delay >= 1) %>%
  arrange(desc(dep_delay))

worst_delays <- flight_weather[c(1:200), c(1:4)]

ggplot(worst_delays, aes(x = temp, y=wind_speed, color = precip)) +
  geom_point()
```
Looking at the plot of the longest 200 departure delays, it appears that cold temperatures and high wind speeds can cause delays, but precipitation is the biggest factor.  



Problem 1 part e
```{r}
# Create a map that sizes each destination airport by number of incoming flights

flights_dest <- flights %>%
  group_by(dest) %>%
  summarise(dest_count = n()) %>%
  inner_join(airports, by = c("dest" = "faa"))

flights_dest %>%
  ggplot(aes(lon, lat, size = dest_count)) +
  borders("state") +
  geom_point(stat="identity", alpha=0.5, color="darkblue") +
  coord_quickmap() +
  xlab("Longitude") +
  ylab( "Latitude") +
  ggtitle("Destination Frequency") +
  labs(size = "Number of incoming flights")
  
```





Problem 2
```{r}
#library(usmap)
#library(ggplot2)

us_states <- map_data("state")
vaccine_by_state <- read.csv("covid19_vaccinations_USA.csv", header = TRUE)

vaccine_by_state %>%
  right_join(us_states, by=c("state" = "region"))

min_by_vacc <- min(vaccine_by_state$Janssen_Doses,
                   vaccine_by_state$Moderna_Doses,
                   vaccine_by_state$Pfizer_Doses)

max_by_vacc <- max(vaccine_by_state$Janssen_Doses,
                   vaccine_by_state$Moderna_Doses,
                   vaccine_by_state$Pfizer_Doses)


# Janssen
#vaccine_by_state %>%
#  ggplot(aes(x=long, y=lat, fill=Janssen_Doses)) +
#  borders("state") +
#  geom_polygon(color = "gray90", size = 0.1) +
#  guides(fill="none") +
#  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
#  theme(axis.line = element_blank(),
#        axis.text = element_blank(),
#        axis.ticks = element_blank(),
#        axis.title = element_blank(),
#        panel.background = element_blank(),
#        panel.border = element_blank(),
#        panel.grid = element_blank())

# I could not get this problem to work even after trying a couple different methods.

```



Problem 3
```{r}

library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
library("dplyr")

file_text <- readLines("Repurposing_Project_Report_Final.txt")
corpus <- Corpus(VectorSource(file_text))
corpus <- corpus %>% 
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)

dtm <- TermDocumentMatrix(corpus)
matrix <- as.matrix(dtm)
sorted_matrix <- sort(rowSums(matrix), decreasing=TRUE)
df <- data.frame(word = names(sorted_matrix), frequency = sorted_matrix)

res <- wordcloud(corpus, min.freq = 1, max.words = 200, colors = brewer.pal(6, name="Dark2"))

```

